---
title: "Task 4 - Naples Reef"
author: "Lucas Boyd"
date: "10/24/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# attaching packages
library(tidyverse)
library(here)
library(janitor)
```

## Overview

Insert a short description here.

## Reading in the data

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp
removal: Urchin size frequency distribution ver 20. Environmental Data
Initiative.
<https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b>.

Summary of variables:

-   date: date urchin recorded

-   site: site along the SB coast (the five sites can be explored here)

-   treatment: indicates experimental kelp treatments:

    -   CONTROL = giant kelp left intact

    -   ANNUAL = giant kelp is "removed...once each winter and allowed
        to re-colonize over the course of the year"

-   size: urchin diameter (cm), measured to the nearest 0.5 cm

-   scientific_name: scientific name

-   common_name: species common name (either red or purple urchin)

```{r}

```
